### 2024-09-23 paper/paper.qmd
Initial Format Completed

---
title: "My title"
subtitle: "My subtitle if needed"
author: 
  - Rohan Alexander
thanks: "Code and data are available at: LINK."
date: today
date-format: long
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)

data <- read_csv(here::here("data/analysis_data/analysis_data.csv"))

```

# Introduction

You can and should cross-reference sections and sub-sections. We use @talia, @opendatatorotno, and @rohan.

The remainder of this paper is structured as follows. @sec-mydatasection

# Data {#sec-mydatasection}

Some of our data is of penguins (@fig-marriage), from @palmerpenguins.

```{r}
#| label: fig-marriage
#| fig-cap: Bills of penguins
#| echo: false

data |> 
  ggplot(aes(x = date, y = marriage_licenses)) +
  geom_point()

```

Talk more about it.

And also planes (@fig-planes). (You can change the height and width, but don't worry about doing that until you have finished every other aspect of the paper - Quarto will try to make it look nice and the defaults usually work well once you have enough text.)

# Discussion

## First discussion point {#sec-first-point}

If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.

## Second discussion point

## Third discussion point

## Weaknesses and next steps

Weaknesses and next steps should also be included.

\newpage

\appendix

# Appendix {-}

# Additional data details

\newpage

# References

### 2024-09-23 scripts/00-simulate_data.R

@@ -1,0 +1,38 @@
+#### Preamble ####
+# Purpose: Simulates data
+# Author: Karen Riani
+# Date: 19 September 2024
+# Contact: karen.riani@mail.utoronto.ca
+# Pre-requisites: None
+# Any other information needed? None
+
+#### Workspace setup ####
+library(tidyverse)
+
+
+#### Simulate data ####
+set.seed(304)
+
+# Define the start and end date
+start_date <- as.Date("2018-01-01")
+end_date <- as.Date("2023-12-31")
+
+# Set the number of random dates you want to generate
+number_of_dates <- 100
+
+data <-
+  tibble(
+    dates = as.Date(
+      runif(
+        n = number_of_dates,
+        min = as.numeric(start_date),
+        max = as.numeric(end_date)
+      ),
+      origin = "1970-01-01"
+    ),
+    number_of_marriage = rpois(n = number_of_dates, lambda = 15)
+  )
+
+
+#### Write_csv
+write_csv(data, file = "data/raw_data/simulated.csv")

### 2024-09-23 scripts/01-download_data.R

@@ -1,0 +1,30 @@
+#### Preamble ####
+# Purpose: Downloads and saves the data from Open Data Toronto
+# Author: Karen Riani
+# Date: 19 September 2024
+# Contact: karen.riani@mail.utoronto.ca
+# Pre-requisites: None
+# Any other information needed? None
+
+#### Workspace setup ####
+library(opendatatoronto)
+library(tidyverse)
+
+
+#### Download data ####
+# get package
+package <- show_package("e28bc818-43d5-43f7-b5d9-bdfb4eda5feb")
+
+# get all resources for this package
+resources <- list_package_resources("e28bc818-43d5-43f7-b5d9-bdfb4eda5feb")
+
+# identify datastore resources; by default, Toronto Open Data sets datastore resource format to CSV for non-geospatial and GeoJSON for geospatial resources
+datastore_resources <- filter(resources, tolower(format) %in% c('csv', 'geojson'))
+
+# load the first datastore resource as a sample
+data <- filter(datastore_resources, row_number()==1) %>% 
+  get_resource()
+
+
+#### Save data ####
+write_csv(data, "data/raw_data/raw_data.csv") 

### 2024-09-23 scripts/02-data_cleaning.R

@@ -1,0 +1,25 @@
+#### Preamble ####
+# Purpose: Cleans the raw marriage data into an analysis dataset
+# Author: Karen Riani
+# Date: 19 September 2024
+# Contact: karen.riani@mail.utoronto.ca
+# Pre-requisites: None
+# Any other information needed? None
+
+#### Workspace setup ####
+library(tidyverse)
+
+#### Clean data ####
+raw_data <- read_csv("data/raw_data/raw_data.csv")
+
+cleaned_data <-
+  raw_data |>
+  janitor::clean_names() |> 
+  separate(col = time_period,
+            into = c("year", "month"),
+            sep = "-") |> 
+  mutate(date = lubridate::ymd(paste(year, month, "01", sep = "-"))
+         )
+  
+#### Save data ####
+write_csv(cleaned_data, "data/analysis_data/analysis_data.csv")

### 2024-09-23 scripts/03-test_data.R

@@ -1,0 +1,26 @@
+#### Preamble ####
+# Purpose: Sanity check of the data
+# Author: Karen Riani
+# Date: 19 September 2024
+# Contact: karen.riani@mail.utoronto.ca
+# Pre-requisites: None
+# Any other information needed? None
+
+#### Workspace setup ####
+library(tidyverse)
+
+
+#### Test data ####
+data <- read_csv("data/raw_data/simulated.csv")
+
+# Test for negative numbers
+data$number_of_marriage |> min() <= 0
+
+# Test for NAs
+all(is.na(data$number_of_marriage))
+
+
+
+
+
+

### 2024-09-24 scripts/01-download_data.R
Added download script for Hate Crime data

@@ -1,30 +1,28 @@
 #### Preamble ####
 # Purpose: Downloads and saves the data from Open Data Toronto
 # Author: Karen Riani
-# Date: 19 September 2024
+# Date: 22 September 2024
 # Contact: karen.riani@mail.utoronto.ca
 # Pre-requisites: None
 # Any other information needed? None
 
-#### Workspace setup ####
 library(opendatatoronto)
 library(tidyverse)
+library(dplyr)
 
-
-#### Download data ####
 # get package
-package <- show_package("e28bc818-43d5-43f7-b5d9-bdfb4eda5feb")
+package <- show_package("hate-crimes-open-data")
+package
 
 # get all resources for this package
-resources <- list_package_resources("e28bc818-43d5-43f7-b5d9-bdfb4eda5feb")
+resources <- list_package_resources("hate-crimes-open-data")
 
 # identify datastore resources; by default, Toronto Open Data sets datastore resource format to CSV for non-geospatial and GeoJSON for geospatial resources
 datastore_resources <- filter(resources, tolower(format) %in% c('csv', 'geojson'))
 
 # load the first datastore resource as a sample
-data <- filter(datastore_resources, row_number()==1) %>% 
-  get_resource()
-
+data <- filter(datastore_resources, row_number()==1) %>% get_resource()
+data
 
 #### Save data ####
 write_csv(data, "data/raw_data/raw_data.csv") 

### 2024-09-24 scripts/02-data_cleaning.R
Mutated race and offence groups

@@ -1,7 +1,7 @@
 #### Preamble ####
 # Purpose: Cleans the raw marriage data into an analysis dataset
 # Author: Karen Riani
-# Date: 19 September 2024
+# Date: 22 September 2024
 # Contact: karen.riani@mail.utoronto.ca
 # Pre-requisites: None
 # Any other information needed? None
@@ -12,14 +12,20 @@ library(tidyverse)
 #### Clean data ####
 raw_data <- read_csv("data/raw_data/raw_data.csv")
 
-cleaned_data <-
-  raw_data |>
-  janitor::clean_names() |> 
-  separate(col = time_period,
-            into = c("year", "month"),
-            sep = "-") |> 
-  mutate(date = lubridate::ymd(paste(year, month, "01", sep = "-"))
-         )
-  
+# Recategorize race and primary offences
+cleaned_data <- data %>%
+  mutate(
+    race_bias_recategorized = case_when(
+      str_detect(RACE_BIAS, "East/Southeast Asian") ~ "East Asian mixed",
+      str_detect(RACE_BIAS, "South Asian") ~ "South Asian mixed",
+      TRUE ~ "All the rest"
+    ),
+    primary_offence_recategorized = case_when(
+      PRIMARY_OFFENCE %in% c("Assault", "Aggravated Assault") ~ "Assault",
+      PRIMARY_OFFENCE %in% c("Mischief Under $5000", "Mischief Over $5000") ~ "Mischief",
+      TRUE ~ "Other"
+    )
+  )
+
 #### Save data ####
 write_csv(cleaned_data, "data/analysis_data/analysis_data.csv")

### 2024-09-24 scripts/00-simulate_data.R
Added initial data simulation

@@ -1,7 +1,7 @@
 #### Preamble ####
 # Purpose: Simulates data
 # Author: Karen Riani
-# Date: 19 September 2024
+# Date: 22 September 2024
 # Contact: karen.riani@mail.utoronto.ca
 # Pre-requisites: None
 # Any other information needed? None
@@ -9,30 +9,19 @@
 #### Workspace setup ####
 library(tidyverse)
 
-
 #### Simulate data ####
-set.seed(304)
+set.seed(42)
 
-# Define the start and end date
-start_date <- as.Date("2018-01-01")
-end_date <- as.Date("2023-12-31")
+# Set number of Records
+number_of_records <- 100
 
-# Set the number of random dates you want to generate
-number_of_dates <- 100
-
-data <-
-  tibble(
-    dates = as.Date(
-      runif(
-        n = number_of_dates,
-        min = as.numeric(start_date),
-        max = as.numeric(end_date)
-      ),
-      origin = "1970-01-01"
-    ),
-    number_of_marriage = rpois(n = number_of_dates, lambda = 15)
-  )
-
+# Simulating data
+simulated_data <- tibble(
+  occurrence_year = sample(2018:2024, number_of_records, replace = TRUE),
+  race_bias = sample(c("East Asian mixed", "South Asian mixed", "Other"), number_of_records, replace = TRUE),
+  primary_offence = sample(c("Assault", "Mischief", "Hate Speech", "Other"), number_of_records, replace = TRUE),
+  location_type = sample(unique_location_type, number_of_records, replace = TRUE)
+)
 
 #### Write_csv
 write_csv(data, file = "data/raw_data/simulated.csv")

### 2024-09-24 scripts/00-simulate_data.R
updated simulation groups

@@ -16,11 +16,13 @@ set.seed(42)
 number_of_records <- 100
 
 # Simulating data
+
+# Simulating data
 simulated_data <- tibble(
-  occurrence_year = sample(2018:2024, number_of_records, replace = TRUE),
-  race_bias = sample(c("East Asian mixed", "South Asian mixed", "Other"), number_of_records, replace = TRUE),
-  primary_offence = sample(c("Assault", "Mischief", "Hate Speech", "Other"), number_of_records, replace = TRUE),
-  location_type = sample(unique_location_type, number_of_records, replace = TRUE)
+  OCCURRENCE_YEAR = sample(2018:2024, number_of_records, replace = TRUE),
+  RACE_BIAS = sample(cleaned_data$RACE_BIAS, number_of_records, replace = TRUE),
+  LOCATION_TYPE = sample(cleaned_data$LOCATION_TYPE, number_of_records, replace = TRUE),
+  OFFENCE_GROUP = sample(cleaned_data$OFFENCE_GROUP, number_of_records, replace = TRUE)
 )
 
 #### Write_csv

### 2024-09-24 scripts/03-test_data.R
Tests added

@@ -1,7 +1,7 @@
 #### Preamble ####
 # Purpose: Sanity check of the data
 # Author: Karen Riani
-# Date: 19 September 2024
+# Date: 22 September 2024
 # Contact: karen.riani@mail.utoronto.ca
 # Pre-requisites: None
 # Any other information needed? None
@@ -9,17 +9,16 @@
 #### Workspace setup ####
 library(tidyverse)
 
-
 #### Test data ####
 data <- read_csv("data/raw_data/simulated.csv")
 
-# Test for negative numbers
-data$number_of_marriage |> min() <= 0
+# Test for NAs in specific categorical columns
+any(is.na(data$RACE_BIAS))  # Check RACE_BIAS for NAs
+any(is.na(data$LOCATION_TYPE))  # Check LOCATION_TYPE for NAs
+any(is.na(data$PRIMARY_OFFENCE))  # Check OFFENSE_GROUP for NAs
 
-# Test for NAs
-all(is.na(data$number_of_marriage))
-
-
+# Test for duplicate rows
+any(duplicated(data))
 
 
 

### 2024-09-24 scripts/02-data_cleaning.R
Group by offence, remove racial grouping

@@ -14,18 +14,15 @@ raw_data <- read_csv("data/raw_data/raw_data.csv")
 
 # Recategorize race and primary offences
 cleaned_data <- data %>%
-  mutate(
-    race_bias_recategorized = case_when(
-      str_detect(RACE_BIAS, "East/Southeast Asian") ~ "East Asian mixed",
-      str_detect(RACE_BIAS, "South Asian") ~ "South Asian mixed",
-      TRUE ~ "All the rest"
-    ),
-    primary_offence_recategorized = case_when(
-      PRIMARY_OFFENCE %in% c("Assault", "Aggravated Assault") ~ "Assault",
-      PRIMARY_OFFENCE %in% c("Mischief Under $5000", "Mischief Over $5000") ~ "Mischief",
-      TRUE ~ "Other"
-    )
-  )
+  mutate(OFFENCE_GROUP = case_when(
+    PRIMARY_OFFENCE %in% c("Assault", "Assault With a Weapon", "Assault Causing Bodily Harm", 
+                           "Aggravated Assault", "Assault Peace Officer", "Sexual Assault", "Murder") ~ "Assault-related Offenses",
+    PRIMARY_OFFENCE %in% c("Mischief Under $5000", "Mischief Interfere With Property", "Mischief To Religious Property, Educational Institutions, Etc.",
+                           "Mischief Over $5000", "Mischief To Data", "Arson") ~ "Mischief and Property Damage",
+    PRIMARY_OFFENCE %in% c("Uttering Threats - Bodily Harm", "Uttering Threats - Property", "Criminal Harassment", "Harassing Communications") ~ "Threats and Harassment",
+    PRIMARY_OFFENCE %in% c("Wilful Promotion of Hatred", "Public Incitement of Hatred", "Advocating Genocide", "Disturbing Religious Worship Or Certain Meetings") ~ "Hate-related Offenses",
+    TRUE ~ "Other Offenses"
+  ))
 
 #### Save data ####
 write_csv(cleaned_data, "data/analysis_data/analysis_data.csv")

### 2024-09-24 paper/paper.qmd
inital paper structure

@@
 ---
-title: "My title"
-subtitle: "My subtitle if needed"
+title: "Torontonian Hate Crime Trends (temp)"
+subtitle: "An Analysis of Race Bias in Hate Crime from 2018 to 2024"
 author: 
-  - Rohan Alexander
-thanks: "Code and data are available at: LINK."
+  - Karen Riani
+thanks: "Code and data are available at: https://github.com/karenrni/Torontonian-Hate-Crime-Trends"
 date: today
 date-format: long
-abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
+abstract: "The abstract answers: 1) what was done, 2) what was found, and 3) why this matters (all at a high level).
+Likely four sentences.
+Abstract must make clear what we learn about the world because of this paper."
 format: pdf
@@
 
+# Table of Contents
+1. [Introduction](#introduction)  
+2. [Data](#data)  
+   2.1. [Raw Data](#raw-data)  
+   2.2. [Cleaned Data](#cleaned-data)  
+   2.3. [Basic Summary Statistics of the Data](#basic-summary-statistics-of-the-data)  
+   2.4. [Discussion of Data Selection](#discussion-of-data-selection)  
+3. [Results](#results)  
+   3.1. 
+   3.2.
+   3.3. 
+4. [Discussion](#discussion)  
+5. [Limitations and Next Steps](#limitations-and-next-steps)  
+6. [References](#references)
+
 # Introduction
 
-You can and should cross-reference sections and sub-sections. We use @talia, @opendatatorotno, and @rohan.
+In recent years, there has been a noticeable increase in hate crimes targeting individuals of East and South Asian descent, particularly in diverse urban areas such as Toronto.
+Thes crimes have garnered significant attention amid the global rise of anti-Asian sentiment, not only in physical attacks but also in the rise of online cyberbullying.
+The initial spike in hate incidents was observed during the COVID-19 pandemic, primarily affecting East Asian and Chinese communities.
+In the years that followed, changes in immigration policies, especially those impacting international students, have led to heightened animosity towards South Asian individuals.
+This ongoing trend shows patterns of racial bias that highlights the need to understand the dynamics of hate crimes, not only to promote public safety but also to encourage greater awareness and solidarity within the community.
 
-The remainder of this paper is structured as follows. @sec-mydatasection
+This paper aims to analyze hate crimes targeting Asian populations in Toronto from 2018 to 2024, focusing specifically on the trends associated with various racial identities, including East and South Asian individuals.
+
+To address this gap, the study employs a
+
+The findings reveal significant trends in the prevalence of hate crimes against Asian individuals, with notable increases during key periods influenced by societal events, such as the COVID-19 pandemic and changing immigration policies.
+
+Understanding the patterns of hate crimes against Asian communities is vital for informing policy interventions and community support initiatives.
+
+The structure of this paper is as follows: Section 2 presents the data sources and methodology, followed by Section 3, which discusses the results and their implications.
+Section 4 concludes with a discussion on the limitations of the study and suggestions for future research.
 
@@
 
-Some of our data is of penguins (@fig-marriage), from @palmerpenguins.
+A sense of the dataset should be communicated to the reader.
+The broader context of the dataset should be discussed.
+All variables should be thoroughly examined and explained.
+Explain if there were similar datasets that could have been used and why they were not.
+If variables were constructed then this should be mentioned, and high-level cleaning aspects of note should be mentioned, but this section should focus on the destination, not the journey.
+It is important to understand what the variables look like by including graphs, and possibly tables, of all observations, along with discussion of those graphs and the other features of these data.
+Summary statistics should also be included, and well as any relationships between the variables.
+If this becomes too detailed, then appendices could be used.
+Basically, for every variable in your dataset that is of interest to your paper there needs to be graphs and explanation and maybe tables.
+A thorough discussion of measurement, relating to the dataset, is provided in the data section.
+Please ensure that you explain how we went from some phenomena in the world that happened to an entry in the dataset that you are interested in.
+
+Seen via (@fig-Asian-Bias-Fig).
 
 ```{r}
-#| label: fig-marriage
-#| fig-cap: Bills of penguins
+#| label: fig-Asian-Bias-Fig
+#| fig-cap: Asian Biased Hate Crimes in Toronto
 #| echo: false
 
-data |> 
-  ggplot(aes(x = date, y = marriage_licenses)) +
-  geom_point()
+# To filter for Asian races
+asian_races <- c("East/Southeast Asian", 
+                 "South Asian", 
+                 "Black, East/Southeast Asian", 
+                 "Black, South Asian", 
+                 "East/Southeast Asian, South Asian, Black", 
+                 "White, East/Southeast Asian",
+                 "Indigenous, East/Southeast Asian")
+
+# Filter for Asian-related data
+asian_related_data <- data %>%
+  filter(RACE_BIAS %in% asian_races)
+
+# Aggregate data by year and race bias
+aggregated_data <- asian_related_data %>%
+  group_by(OCCURRENCE_YEAR, RACE_BIAS) %>%
+  summarise(incident_count = n(), .groups = 'drop')
+
+# Create a plot
+ggplot(aggregated_data, aes(x = OCCURRENCE_YEAR, y = incident_count, fill = RACE_BIAS)) +
+  geom_bar(stat = "identity", position = "dodge") +
+  labs(title = "Hate Crimes Involving Asian Races in Toronto (2018-2024)",
+       x = "Year",
+       y = "Number of Incidents",
+       fill = "Race Bias") +
+  theme_minimal() +
+  theme(axis.text.x = element_text(angle = 45, hjust = 1))
 
@@
 
-Talk more about it.
+```{r}
+#| label: fig-Race-Bias-Fig
+#| fig-cap: Race Biased Hate Crimes in Toronto
+#| echo: false
 
-And also planes (@fig-planes).
-(You can change the height and width, but don't worry about doing that until you have finished every other aspect of the paper - Quarto will try to make it look nice and the defaults usually work well once you have enough text.)
+hate_crime_counts <- data %>%
+  group_by(OCCURRENCE_YEAR, RACE_BIAS) %>%
+  summarise(count = n(), .groups = 'drop')
+
+# Plotting
+ggplot(hate_crime_counts, aes(x = OCCURRENCE_YEAR, y = count, fill = RACE_BIAS)) +
+  geom_bar(stat = "identity", position = "dodge") +
+  labs(title = "Hate Crimes in Toronto (2018-2024)",
+       subtitle = "Count of Hate Crimes by Race Bias",
+       x = "Year",
+       y = "Number of Hate Crimes") +
+  theme_minimal() +
+  theme(axis.text.x = element_text(angle = 45, hjust = 1))
+```
+
+### Raw Data
+source, structure,
+
+### Cleaned Data
+recategorization of variables.
+
+### Basic Summary Statistics of the Data
+
+### Discussion of Data Selection
+Discuss the rationale behind the selection of the data used for analysis and any criteria applied.
+
+## Results
 
@@
 
-If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.
-
 ## Second discussion point
@@
 
-## Weaknesses and next steps
+# Limitations and Next Steps
 
-Weaknesses and next steps should also be included.
-
-\newpage
-
-\appendix
-
-# Appendix {-}
-
-# Additional data details
+- accuracy and reliability of reports
+- dataset includes only all verified Hate Crime occurrences investigated by the Hate Crime Unit by 
+- only those confirmed and reported (date since 2018) 
+- does not include occurrences that have been deemed unfounded or classified as hate incidents
+- includes interaction between multi-biased hate crimes, where motivation lacks clarity 
+  - i.e. Age, Mental or Physical Disability, Race, Ethnicity, Language, Religion, Sexual Orientation, Gender and Other Similar Factor
+- should contine to analyse with multi level bias crimes in mind
+- data is provided at the offence and/or occurrence level-- one occurrence may have multi-bias categories associated to the victim used to categorize the hate crime
+- may overlook qualitative aspects of hate crimes, such as the socio-cultural context of incidents
 
@@
 # References
+
+Gelfand, S. (2022). opendatatoronto: Access the City of Toronto Open Data Portal (R package version 0.1.5). Retrieved from https://CRAN.R-project.org/package=opendatatoronto
+
+R Core Team. (2023). R: A Language and Environment for Statistical Computing. Vienna, Austria: R Foundation for Statistical Computing. Retrieved from https://www.R-project.org/

### 2024-09-24 scripts/00-simulate_data.R
updates for clarity

@@ -13,9 +13,7 @@ library(tidyverse)
 set.seed(42)
 
 # Set number of Records
-number_of_records <- 100
-
-# Simulating data
+number_of_records <- 300
 
 # Simulating data
 simulated_data <- tibble(

### 2024-09-24 scripts/02-data_cleaning.R

@@ -24,5 +24,9 @@ cleaned_data <- data %>%
     TRUE ~ "Other Offenses"
   ))
 
+# Remove rows without racial bias 
+cleaned_data <- cleaned_data %>%
+  filter(RACE_BIAS != "None")
+
 #### Save data ####
 write_csv(cleaned_data, "data/analysis_data/analysis_data.csv")

### 2024-09-27 scripts/02-data_cleaning.R
Additional cleaning

@@ -1,7 +1,7 @@
 #### Preamble ####
 # Purpose: Cleans the raw marriage data into an analysis dataset
 # Author: Karen Riani
-# Date: 22 September 2024
+# Date: 27 September 2024
 # Contact: karen.riani@mail.utoronto.ca
 # Pre-requisites: None
 # Any other information needed? None
@@ -12,8 +12,21 @@ library(tidyverse)
 #### Clean data ####
 raw_data <- read_csv("data/raw_data/raw_data.csv")
 
-# Recategorize race and primary offences
-cleaned_data <- data %>%
+# Ensure correct date format
+raw_data$OCCURRENCE_DATE <- as.Date(raw_data$OCCURRENCE_DATE, format = "%Y-%m-%d")
+
+# Check duplicates based on _id
+duplicate_rows <- raw_data[duplicated(raw_data$EVENT_UNIQUE_ID), ]
+
+# Remove duplicates
+data_cleaned <- raw_data[!duplicated(raw_data$EVENT_UNIQUE_ID), ]
+
+# Rename "Other" to "Unknown" in the RACE_BIAS column
+data <- data %>%
+  mutate(RACE_BIAS = recode(RACE_BIAS, "Other race not mentioned" = "Unknown"))
+
+# Consolidate primary offence types to capture broader categories
+cleaned_data <- raw_data %>%
   mutate(OFFENCE_GROUP = case_when(
     PRIMARY_OFFENCE %in% c("Assault", "Assault With a Weapon", "Assault Causing Bodily Harm", 
                            "Aggravated Assault", "Assault Peace Officer", "Sexual Assault", "Murder") ~ "Assault-related Offenses",

### 2024-09-27 scripts/00-simulate_data.R
minor updates

@@ -15,7 +15,7 @@ set.seed(42)
 # Set number of Records
 number_of_records <- 300
 
-# Simulating data
+# Simulating data with desired variables
 simulated_data <- tibble(
   OCCURRENCE_YEAR = sample(2018:2024, number_of_records, replace = TRUE),
   RACE_BIAS = sample(cleaned_data$RACE_BIAS, number_of_records, replace = TRUE),

### 2024-09-27 paper/paper.qmd
Minor fixes

--- 2024-09-23 paper/paper.qmd
@@
 ---
-title: "My title"
-subtitle: "My subtitle if needed"
+title: "Torontonian Hate Crime Trends (temp)"
+subtitle: "An Analysis of Race Bias in Hate Crime from 2018 to 2023"
 author: 
-  - Rohan Alexander
-thanks: "Code and data are available at: LINK."
+  - Karen Riani
+thanks: "Code and data are available at: https://github.com/karenrni/Torontonian-Hate-Crime-Trends"
 date: today
 date-format: long
-abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
+abstract: "The abstract answers: 1) what was done, 2) what was found, and 3) why this matters (all at a high level).
+Likely four sentences.
+Abstract must make clear what we learn about the world because of this paper."
 format: pdf
@@
 library(tidyverse)
+library(dplyr)
+library(knitr)
 
@@
 
-You can and should cross-reference sections and sub-sections. We use @talia, @opendatatorotno, and @rohan.
+In recent years, there has been a noticeable increase in hate crimes targeting individuals of East and South Asian descent, particularly in diverse urban areas such as Toronto.
+These crimes have garnered significant attention amid the global rise of anti-Asian sentiment, not only in physical attacks but also in the rise of online cyberbullying.
+The initial spike in hate incidents was observed during the COVID-19 pandemic, primarily affecting East Asian and Chinese communities.
+Furthermore, changes in immigration policies, especially those impacting international students, have led to heightened animosity towards South Asian individuals.
+This ongoing trend shows patterns of racial bias that highlight the need to understand the dynamics of hate crimes, not only to promote public safety but also to encourage greater awareness and solidarity within the community.
 
-The remainder of this paper is structured as follows. @sec-mydatasection
+Amidst the rise in Asian hate in the month of March 2020, the Canadian government began more actively working with Asian communities to handle increased anti-Asian sentiment heightened by COVID-19.
+Several policies were recommended by these communities, and the Canadian government responded in May 2021 by developing a definition of anti-Asian racism, which was added to Canada’s Anti-Racism Strategy (Government of Canada, 2023).
+
+The Canadian Race Relations Foundation (CRRF) and Public Safety Canada additionally organized several engagement initiatives and published statements for Asian communities, i.e.
+Asian and Sikh Heritage Month and The Bias, Sensitivity, Diversity and Identity Symposium on Islamophobia and Anti-Asian Hate.
+These efforts provided a platform for Asian Canadians to voice their experiences and challenges.
+Expert discussions further highlighted the impact of national security policies on Muslim and Asian communities amid rising hate crimes.
+
+The 2020 policy changes allowing international students to study online while preserving work permit eligibility helped mitigate disruptions but did not address the broader rise in anti-Asian racism (Government of Canada, 2020).
+Social media fueled this sentiment, leading to increased hate crimes against South Asian students in public and online spaces, especially in Toronto (The Link, 2022).
+
+Despite these policy responses, Asian communities continue to face systemic discrimination.
+This paper examines hate crimes against Asian populations in Toronto from 2018 to 2023, with a focus on how socio-political developments, like the pandemic and policy shifts, have shaped these trends.
+While acknowledging broader racial disparities affecting Black communities, this paper narrows in on anti-Asian bias, aiming to inform more targeted interventions and support strategies for affected groups.
+
+The following structure of this paper is as follows: Section 2 presents the data sources and methodology, followed by Section 3, which discusses the results and their implications.
+Section 4 concludes with a discussion of the limitations of the study and suggestions for future research.
 
@@
 
-Some of our data is of penguins (@fig-marriage), from @palmerpenguins.
+The data used in this paper is sourced from the City of Toronto's Open Data Portal, accessed via the opendatatoronto package (Gelfand 2022).
+The dataset, Hate Crimes Open Data (City of Toronto 2024), includes verified hate crime occurrences reported to the Toronto Police since 2018.
+Analysis was conducted using R (R Core Team 2023), with key packages dplyr (Wickham et al.
+2023), tibble (Müller et al.
+2023), tidyverse (Wickam et al.
+2019), and ggplot2 (Wickham 2016).
+Report generation was done with knitr (Xie 2023).
+
+## Hate Crime Data
+
+Hate crime data provided by the City of Toronto’s Hate Crime Unit includes any reported hate crime occurrence beginning from January 2018.
+The features include identifiers like offence and report details (ID, year, date, time), police division and location, along with biases related to age, disability, race, ethnicity, language, religion, sexual orientation, and gender, as well as neighbourhood identifiers and whether an arrest was made.
+
+The set includes 1350 hate crimes from January 2018 to most recently, December 2023.
+It is annually refreshed, with the latest occurring on September 9th, 2024.
+As for quality, the data has an overall score of 96% grade of gold, and above 97% completeness, usability and freshness.
 
 ```{r}
-#| label: fig-marriage
-#| fig-cap: Bills of penguins
+#| label: fig-Race-Bias-Fig
+#| fig-cap: Race Biased Hate Crimes in Toronto
 #| echo: false
 
-data |> 
-  ggplot(aes(x = date, y = marriage_licenses)) +
-  geom_point()
+#Count total hate crimes
+hate_crime_counts <- data %>%
+  group_by(OCCURRENCE_YEAR, RACE_BIAS) %>%
+  summarise(count = n(), .groups = 'drop')
+
+## Plotting ##
+
+# Find the top 4 races with the most hate crimes per year
+top_per_year <- hate_crime_counts %>%
+  group_by(OCCURRENCE_YEAR) %>%
+  top_n(4, wt = count)
+
+# Plotting and detailing
+ggplot(top_per_year, aes(x = OCCURRENCE_YEAR, y = count, fill = RACE_BIAS)) +
+  geom_bar(stat = "identity", position = "dodge") +
+  labs(title = "Top 4 Hate Crimes by Race Bias (2018-2023)",
+       x = "Year",
+       y = "Number of Hate Crimes",
+       fill = "Race Bias") +
+  theme_minimal() +
+  theme(axis.text.x = element_text(angle = 45, hjust = 1),
+        legend.title = element_text(size = 10),
+        legend.text = element_text(size = 8))      
+```
+
+Seen via (@fig-Asian-Bias-Fig).
+
+```{r}
+#| label: fig-Asian-Bias-Fig
+#| fig-cap: Asian Biased Hate Crimes in Toronto
+#| echo: false
+
+# To filter for Asian races
+asian_races <- c("East/Southeast Asian", 
+                 "South Asian", 
+                 "Black, East/Southeast Asian", 
+                 "Black, South Asian", 
+                 "East/Southeast Asian, South Asian, Black", 
+                 "White, East/Southeast Asian",
+                 "Indigenous, East/Southeast Asian")
+
+# Filter for Asian-related data
+asian_data <- data %>%
+  filter(RACE_BIAS %in% asian_races)
+
+# Aggregate data by year and race bias
+aggregated_data <- asian_data %>%
+  group_by(OCCURRENCE_YEAR, RACE_BIAS) %>%
+  summarise(incident_count = n(), .groups = 'drop')
+
+# Create a plot
+ggplot(aggregated_data, aes(x = OCCURRENCE_YEAR, y = incident_count, fill = RACE_BIAS)) +
+  geom_bar(stat = "identity", position = "dodge") +
+  labs(title = "Hate Crimes Involving Asian Races in Toronto (2018-2024)",
+       x = "Year",
+       y = "Number of Incidents",
+       fill = "Race Bias") +
+  theme_minimal() +
+  theme(axis.text.x = element_text(angle = 45, hjust = 1),
+        legend.text = element_text(size = 7))
 
@@
 
-Talk more about it.
+```{r}
+#| label: fig-Asian-Bias-Summary
+#| fig-cap: Summary of Asian Biased Hate Crimes in Toronto
+#| echo: false
 
-And also planes (@fig-planes).
-(You can change the height and width, but don't worry about doing that until you have finished every other aspect of the paper - Quarto will try to make it look nice and the defaults usually work well once you have enough text.)
+# Generate summary statistics for Asian bias data 
+summary <- asian_data %>%
+  group_by(OCCURRENCE_YEAR) %>%
+  summarise(
+    `Incident Count` = n(),
+    `Race Biases` = n_distinct(RACE_BIAS),
+    `Location Types` = n_distinct(LOCATION_TYPE),
+    `Offence Groups` = n_distinct(OFFENCE_GROUP)
+  ) %>%
+  rename(`Year` = OCCURRENCE_YEAR)
+
+# Display the summary as a table using kable
+kable(summary, caption = "Summary of Asian Biased Hate Crimes in Toronto")
+
+```
+
+```{r}
+#| label: fig-East-Asn-Summary
+#| fig-cap: Summary of increase in hate crimes against East/Southeast Asians
+#| echo: false
+
+# Group by OCCURRENCE_YEAR and RACE_BIAS to summarize the counts
+Easian_crimes <- data %>%
+  filter(RACE_BIAS %in% c("East/Southeast Asian")) %>%
+  group_by(OCCURRENCE_YEAR, RACE_BIAS) %>%
+  summarise(count = n(), .groups = 'drop')
+
+# Calculate year-on-year increase
+Easian_crimes <- Easian_crimes %>%
+  arrange(RACE_BIAS, OCCURRENCE_YEAR) %>%
+  group_by(RACE_BIAS) %>%
+  mutate(increase = (count - lag(count)) / lag(count) * 100)  # Percentage increase
+
+# Display the result as a table
+Easian_crimes %>%
+  kable(caption = "Year-on-Year Increase in Hate Crimes Against Asians")
+
+```
+
+```{r}
+#| label: fig-South-Asn-Summary
+#| fig-cap: Summary of increase in hate crimes against South Asians
+#| echo: false
+
+# Group by OCCURRENCE_YEAR and RACE_BIAS to summarize the counts
+Sasian_crimes <- data %>%
+  filter(RACE_BIAS %in% c("South Asian")) %>%
+  group_by(OCCURRENCE_YEAR, RACE_BIAS) %>%
+  summarise(count = n(), .groups = 'drop')
+
+# Calculate year-on-year increase
+Sasian_crimes <- Sasian_crimes %>%
+  arrange(RACE_BIAS, OCCURRENCE_YEAR) %>%
+  group_by(RACE_BIAS) %>%
+  mutate(increase = (count - lag(count)) / lag(count) * 100)  # Percentage increase
+
+# Display the result as a table
+Sasian_crimes %>%
+  kable(caption = "Year-on-Year Increase in Hate Crimes Against Asians")
+
+```
+### Raw Data
+source, structure,
+
+### Cleaned Data
+recategorization of variables.
+
+### Basic Summary Statistics of the Data
+
+### Discussion of Data Selection
+Discuss the rationale behind the selection of the data used for analysis and any criteria applied.
+
+## Results
 
@@
 # References
+
+City of Toronto. 2024. Hate Crimes Open Data. Published by Policy, Planning, Finance & Administration. Retrieved from https://www.toronto.ca/open-data-catalogue/hate-crimes-open-data.
+
+Gelfand, Sharla. 2022. Opendatatoronto: Access the City of Toronto Open Data Portal. Retrieved from https://CRAN.R-project.org/package=opendatatoronto.
+
+Government of Canada.
+2020.
+Measures to Support International Students During the COVID-19 Pandemic.
+Immigration, Refugees and Citizenship Canada.
+Retrieved from https://www.canada.ca/en/immigration-refugees-citizenship/services/coronavirus-covid19.html.
+
+Government of Canada.
+2023.
+Annual Report on the Operation of the Canadian Multiculturalism Act 2021-2022.
+Retrieved from https://www.canada.ca/en/canadian-heritage/corporate/publications/plans-reports/annual-report-canadian-multiculturalism-act-2021-2022.html.
+
+Hadley Wickham, Lionel Henry, Romain François. 2023. Dplyr: A Grammar of Data Manipulation. Retrieved from https://CRAN.R-project.org/package=dplyr.
+
+Kirill Müller, Romain Francois, Hadley Wickham. 2023. Tibble: Simple Data Frames. Retrieved from https://CRAN.R-project.org/package=tibble.
+
+The Link. 2022. Canada Has a Rampant South Asian Racism Problem. The Link Newspaper. Retrieved from https://thelinknewspaper.ca/article/canada-has-a-rampant-south-asian-racism-problem.
+
+R Core Team. 2023. R: A Language and Environment for Statistical Computing. Vienna, Austria: R Foundation for Statistical Computing. Retrieved from https://www.R-project.org/.
+
+Wickham, Hadley. 2016. Ggplot2: Elegant Graphics for Data Analysis. Springer-Verlag New York. Retrieved from https://ggplot2.tidyverse.org.
+
+Wickham, Hadley, Mara Averick, Jennifer Bryan, Winston Chang, Lucy D’Agostino McGowan, Romain François, Garrett Grolemund, et al.
+2019.
+“Welcome to the Tidyverse.” Journal of Open Source Software 4 (43): 1686.
+Retrieved from https://doi.org/10.21105/joss.01686.
+
+Xie, Yihui. 2023. Knitr: A General-Purpose Package for Dynamic Report Generation in R. Retrieved from https://yihui.org/knitr/.

### 2024-09-28 paper/paper.qmd
minor updates

@@
 The features include identifiers like offence and report details (ID, year, date, time), police division and location, along with biases related to age, disability, race, ethnicity, language, religion, sexual orientation, and gender, as well as neighbourhood identifiers and whether an arrest was made.
+These biases are defined as “crimes committed on the basis of the Suspect’s perception of the Victim’s race”, (City of Toronto 2024).
 
@@
 As for quality, the data has an overall score of 96% grade of gold, and above 97% completeness, usability and freshness.
+The initial set of data included all races and mixes, with the column “Other race not mentioned” renamed to “Unknown” and all crimes not mainly associated with race bias removed– any row containing “None” under RACE_BIAS.
+
+The columns included in the analysis are Primary Offences (Regrouped), Race Bias and Occurance Year.
+Due to the analysis’ focus on Asian Bias hate crimes, races of only East, Southeast and South Asian descent were included.
+Furthermore, primary offences were regrouped into more general overarching categories including, Assault-Related, Threats and Harassment, Mischief & Property Damage, Hate-Related and Others.
+Only the occurrence year was included as opposed to full dates, and reported dates were used to focus on high-level trends.
+Additionally, the dataset did not contain information on other socio-economic factors related to location, such as mean income levels and ethnic demographics, thus locations were excluded.
+Furthermore, location type was recategorized into more general categories including public, private, government, casual and other places.
+
+# Results
+
+## Overview of Hate Crimes by Race
+We are primarily interested in hate crimes targeting Asian races, as this aligns with our focus on anti-Asian sentiment.
+However, it is also important to examine hate crimes across all racial categories to provide a comprehensive understanding of the trends.
+In @fig-Race-Bias-Fig, the graph hows a consistent increase in hate crimes targeting the Black community.
+Notably, hate crimes against Black individuals have been significantly higher than those targeting other racial groups throughout the observed period.
 
@@
 
-Seen via (@fig-Asian-Bias-Fig).
+## Hate Crimes Against Asian Communities
+
+@fig-Asian-Bias-Fig highlights the trend in hate crimes targeting all Asian racial groups, with a year-by-year breakdown.
+This visualization demonstrates a rise in hate crimes against Asian communities, especially during the years corresponding to the COVID-19 pandemic.
+Although the overall trend for Asian hate crimes has fluctuated, the data shows an alarming increase starting in 2020.
 
@@
 ```
+
+## Summary Statistics: Asian and South Asian Communities
+
+@fig-Asian-Bias-Summary presents summary statistics for hate crimes specifically targeting Asian racial groups.
+fig-East-Asn-Summary and fig-South-Asn-Summary show the increase in the 2 major Asian race groups being East/Southeast Asians and South Asians, respectively.
+The data reveals a steady increase in incidents against South Asians from 2018 to 2023, with notable spikes in crimes against East As Asians in particular.
+The increase in Asian biased crimes also show an increase in different offences and location of offence occurrences.
 
@@
 Easian_crimes %>%
-  kable(caption = "Year-on-Year Increase in Hate Crimes Against Asians")
+  kable(caption = "Year-on-Year Increase in Hate Crimes Against East/ Southeast Asians")
 
@@
 Sasian_crimes %>%
-  kable(caption = "Year-on-Year Increase in Hate Crimes Against Asians")
+  kable(caption = "Year-on-Year Increase in Hate Crimes Against South Asians")
 
 ```
-### Raw Data
-source, structure,
 
-### Cleaned Data
-recategorization of variables.
+```{r}
+#| label: fig-Asian-Offence-Types
+#| fig-cap: Type of Offences in Asian Biased Hate Crimes
+#| echo: false
 
-### Basic Summary Statistics of the Data
+# Group by OCCURRENCE_YEAR and OFFENCE_GROUP for Asian hate crimes
+asian_offense_data <- asian_data %>%
+  group_by(OCCURRENCE_YEAR, OFFENCE_GROUP) %>%
+  summarise(incident_count = n(), .groups = 'drop')
 
-### Discussion of Data Selection
-Discuss the rationale behind the selection of the data used for analysis and any criteria applied.
+# Plotting offense types
+ggplot(asian_offense_data, aes(x = OCCURRENCE_YEAR, y = incident_count, fill = OFFENCE_GROUP)) +
+  geom_bar(stat = "identity", position = "dodge") +
+  labs(title = "Type of Offences in Asian Biased Hate Crimes (2018-2023)",
+       x = "Year",
+       y = "Number of Incidents",
+       fill = "Offence Type") +
+  theme_minimal() +
+  theme(axis.text.x = element_text(angle = 45, hjust = 1),
+        legend.text = element_text(size = 7))
+```
 
-## Results
+```{r}
+#| label: fig-Asian-Location-Types
+#| fig-cap: Location Types in Asian Biased Hate Crimes
+#| echo: false
+
+# Group by OCCURRENCE_YEAR and LOCATION_TYPE for Asian hate crimes
+asian_location_data <- asian_data %>%
+  group_by(OCCURRENCE_YEAR, LOCATION_GROUP) %>%
+  summarise(incident_count = n(), .groups = 'drop')
+
+# Plotting location types
+ggplot(asian_location_data, aes(x = OCCURRENCE_YEAR, y = incident_count, fill = LOCATION_GROUP)) +
+  geom_bar(stat = "identity", position = "dodge") +
+  labs(title = "Location Types in Asian Biased Hate Crimes (2018-2023)",
+       x = "Year",
+       y = "Number of Incidents",
+       fill = "Location Type") +
+  theme_minimal() +
+  theme(axis.text.x = element_text(angle = 45, hjust = 1),
+        legend.text = element_text(size = 7))
+
+```
 
@@
 
-## First discussion point {#sec-first-point}
+Based on the above figures, there is evidence that
 
-## Second discussion point
-
-## Third discussion point
+The surge in hate crimes against East/Southeast Asians during 2020 and 2021 correlates strongly with the social and political context surrounding the COVID-19 pandemic.
+As the virus's origins were linked to Asia, widespread misinformation and xenophobia fueled anti-Asian sentiment.
+The sharp rise in hate crimes during this period is indicative of a broader global trend of increased racial targeting and discrimination.
+While 2022 and 2023 saw a reduction in incidents, the numbers remain higher than pre-pandemic levels, suggesting that the effects of this wave of racism may have long-term impacts.
+In contrast, the steady rise in hate crimes against South Asians points to a broader, underlying trend of racial bias that extends beyond the pandemic.
+Anti-immigrant sentiment, exacerbated by policies affecting international students and immigration more generally, likely played a role in these trends.
+Despite government interventions aimed at supporting immigrant communities during the pandemic (Government of Canada, 2020), the persistent hostility faced by South Asians reflects the ongoing challenges of systemic racism in Canadian society.
+Meanwhile, the consistent levels of anti-Black hate crimes highlight ongoing systemic racism.
+Although hate crimes against Asians have decreased since their peak, the persistence of incidents reflects the long-term impact of these biases, necessitating continued policy focus on combating racial discrimination.
+Both the rise and persistence of these hate crimes underscore the importance of targeted policy responses.
+While public health measures and immigration policies were necessary during the pandemic, they were insufficient in addressing the racial tensions that surfaced.
+There is a clear need for long-term strategies aimed at combating racial bias and xenophobia, especially as it pertains to both East/Southeast and South Asian communities in Canada.
 
@@
 
-- accuracy and reliability of reports
-- dataset includes only all verified Hate Crime occurrences investigated by the Hate Crime Unit by 
-- only those confirmed and reported (date since 2018) 
-- does not include occurrences that have been deemed unfounded or classified as hate incidents
-- includes interaction between multi-biased hate crimes, where motivation lacks clarity 
-  - i.e. Age, Mental or Physical Disability, Race, Ethnicity, Language, Religion, Sexual Orientation, Gender and Other Similar Factor
-- should contine to analyse with multi level bias crimes in mind
-- data is provided at the offence and/or occurrence level-- one occurrence may have multi-bias categories associated to the victim used to categorize the hate crime
-- may overlook qualitative aspects of hate crimes, such as the socio-cultural context of incidents
+There are several limitations to this analysis that should be considered when interpreting the results.
+The accuracy and reliability of reports may be affected by underreporting or misclassification, as the dataset includes only reported and verified hate crimes by the Hate crime Unit.
+Many incidents may go unreported or are excluded occurrences classified as unfounded or those identified as hate incidents.
+As a result, the data may not fully reflect the actual or full range of hate crimes occurring in the GTA community.
+The limited focus may omit significant portions of incidents not classified as hate crimes which still reflect underlying societal tensions.
+
+Some incidents in involve multi-bias hate crimes, where multiple motivations such as race, age, disability, and religion can overlap.
+This can make it difficult to isolate patterns of specific bias categories, as one crime may reflect more than one motivation.
+Understanding the full nature of these crimes becomes more complex when multiple biases are at play.
+
+The dataset lacks important qualitative insights into the socio-cultural context of hate crimes.
+Factors such as mean area income, community dynamics, media influence, and social tensions that may drive hate crimes are not reflected in the dataset, which limits a full understanding of the underlying causes and conditions of these incidents.
+
+In future research, there should be continued focus on multi-bias hate crimes in greater detail, perhaps disaggregating incidents with multiple biases to better understand intersections between race and other categories like age and religion.
+This will help clarify how different forms of discrimination intersect in hate crime occurrences.
+Additionally, it would be beneficial to incorporate qualitative data into future studies.
+Interviews with victims, community feedback, and analysis of media or social media trends could provide richer insights into the social and cultural factors that drive hate crimes.
+It may also be valuable to collaborate with community organizations to gather additional data on underreported hate crimes.
+Partnering with other groups, especially Asian based communities, may help fill gaps in official reports and ensure that marginalized voices are included.
+
+Quantitative additions regarding income levels may also help in understanding the effects of major socio-economic factors on hate crime.
+Expanding the dataset to include a longer time frame beyond 2018 and considering incidents classified as hate-related but not fully investigated as hate crimes could provide a broader perspective on trends over time.
+Such an expansion would allow for more comprehensive insights into the long-term patterns of hate incidents in Toronto.
+
+Furthermore, continued exploration of offense and location trends can be vital Analyzing the types of offenses and the locations where hate crimes against specific groups, particularly Black, East/Southeast Asians and South Asians, occur can help identify hotspots and recurring incidents.
+This information could be used to inform targeted interventions and policy efforts aimed at reducing hate crime occurrences in these areas.
+
+By addressing these limitations and building on the current analysis, future research can provide a deeper and more complex understanding of hate crimes and motivations in Toronto and support more effective strategies for combating them.
 

### 2024-09-28 scripts/00-simulate_data.R

@@ -19,7 +19,7 @@ number_of_records <- 300
 simulated_data <- tibble(
   OCCURRENCE_YEAR = sample(2018:2024, number_of_records, replace = TRUE),
   RACE_BIAS = sample(cleaned_data$RACE_BIAS, number_of_records, replace = TRUE),
-  LOCATION_TYPE = sample(cleaned_data$LOCATION_TYPE, number_of_records, replace = TRUE),
+  LOCATION_GROUP = sample(cleaned_data$LOCATION_GROUP, number_of_records, replace = TRUE),
   OFFENCE_GROUP = sample(cleaned_data$OFFENCE_GROUP, number_of_records, replace = TRUE)
 )
 

### 2024-09-28 scripts/02-data_cleaning.R

@@ -41,5 +41,20 @@ cleaned_data <- raw_data %>%
 cleaned_data <- cleaned_data %>%
   filter(RACE_BIAS != "None")
 
+# Regroup location type to more general categories
+cleaned_data <- cleaned_data %>%
+  mutate(LOCATION_GROUP = case_when(
+    LOCATION_TYPE %in% c("Public Transportation", "Open Area, Park or Parking Lot", "Streets/Roadways/Highway", 
+                         "Business/Retail", "Educational Institution (Universities, Colleges, Schools, etc.)",
+                         "Religious Place of Worship/Cultural Centre") ~ "Public",
+    LOCATION_TYPE %in% c("House (Townhouse, Retirement Home, Garage, Vehicle, Cottage)", 
+                         "Apartment Building (Condo, Retirement Buidling, etc.)") ~ "Private",
+    LOCATION_TYPE %in% c("Government Building (Courthouse, Museums, Parliament Building, etc.)", 
+                         "Government Building") ~ "Government",
+    LOCATION_TYPE %in% c("Other Commercial / Corporate Places", "Non-Commercial/Non for Profit", 
+                         "Medical Facility (Hospitals, Long-term Care, etc.)") ~ "Corporate",
+    TRUE ~ "Other"  # Catch any NA or other values
+  ))
+
 #### Save data ####
 write_csv(cleaned_data, "data/analysis_data/analysis_data.csv")

### 2024-09-28 paper/paper.qmd
finalized paper

@@
 ---
-title: "Torontonian Hate Crime Trends (temp)"
-subtitle: "An Analysis of Race Bias in Hate Crime from 2018 to 2023"
+title: "Asian Biased Hate Crime Trends in Toronto"
+subtitle: "An Analysis of the increased Asian Race Bias in Hate Crimes from 2018 to 2023"
 author: 
@@
 date-format: long
-abstract: "The abstract answers: 1) what was done, 2) what was found, and 3) why this matters (all at a high level).
-Likely four sentences.
-Abstract must make clear what we learn about the world because of this paper."
+abstract: "In this study, we analyzed hate crime trends in Toronto from 2018 to 2023, focusing on racial bias against East, Southeast, and South Asian communities.
+Most incidents occurred in public settings, such as businesses or on the street, with offences primarily involving assault, threats, or mischief.
+Results aligned with major global events and found that hate crimes against East/Southeast Asians surged during the COVID-19 pandemic, while incidents targeting South Asian communities steadily increased beginning from COVID-19.
+These findings highlight the role of public spaces in hate crime incidents and emphasize the need for targeted interventions to address both short-term spikes and persistent racial bias in Toronto."
 format: pdf
@@
 Amidst the rise in Asian hate in the month of March 2020, the Canadian government began more actively working with Asian communities to handle increased anti-Asian sentiment heightened by COVID-19.
-Several policies were recommended by these communities, and the Canadian government responded in May 2021 by developing a definition of anti-Asian racism, which was added to Canada’s Anti-Racism Strategy (Government of Canada, 2023).
+Several policies were recommended by these communities, and the Canadian government responded in May 2021 by developing a definition of anti-Asian racism, which was added to Canada’s Anti-Racism Strategy (@gov_canada_2023_multiculturalism).
 
@@
 
-The 2020 policy changes allowing international students to study online while preserving work permit eligibility helped mitigate disruptions but did not address the broader rise in anti-Asian racism (Government of Canada, 2020).
-Social media fueled this sentiment, leading to increased hate crimes against South Asian students in public and online spaces, especially in Toronto (The Link, 2022).
+The 2020 policy changes allowing international students to study online while preserving work permit eligibility helped mitigate disruptions but did not address the broader rise in anti-Asian racism (@gov_canada_2020_students).
+Social media fueled this sentiment, leading to increased hate crimes against South Asian students in public and online spaces, especially in Toronto (@the_link_2022).
 
@@
 
-The following structure of this paper is as follows: Section 2 presents the data sources and methodology, followed by Section 3, which discusses the results and their implications.
+The following structure of this paper is as follows: Section @sec-mydatasection presents the data sources and methodology, followed by Section 3, which discusses the results and their implications.
 Section 4 concludes with a discussion of the limitations of the study and suggestions for future research.
@@
 
-The data used in this paper is sourced from the City of Toronto's Open Data Portal, accessed via the opendatatoronto package (Gelfand 2022).
-The dataset, Hate Crimes Open Data (City of Toronto 2024), includes verified hate crime occurrences reported to the Toronto Police since 2018.
-Analysis was conducted using R (R Core Team 2023), with key packages dplyr (Wickham et al.
-2023), tibble (Müller et al.
-2023), tidyverse (Wickam et al.
-2019), and ggplot2 (Wickham 2016).
-Report generation was done with knitr (Xie 2023).
+The data used in this paper is sourced from the City of Toronto's Open Data Portal, accessed via the opendatatoronto package (@opendatatoronto).
+The dataset, Hate Crimes Open Data (@toronto_hatecrimes_2024), includes verified hate crime occurrences reported to the Toronto Police since 2018.
+Analysis was conducted using R (@r_core_team_2023), with key packages such as dplyr (@dplyr_2023), tibble (@tibble_2023), tidyverse (@tidyverse_welcome_2019), and ggplot2 (@ggplot2_wickham_2016).
+Report generation was done with knitr (@knitr_xie_2023).
 
@@
 The features include identifiers like offence and report details (ID, year, date, time), police division and location, along with biases related to age, disability, race, ethnicity, language, religion, sexual orientation, and gender, as well as neighbourhood identifiers and whether an arrest was made.
-These biases are defined as “crimes committed on the basis of the Suspect’s perception of the Victim’s race”, (City of Toronto 2024).
+These biases are defined as “crimes committed on the basis of the Suspect’s perception of the Victim’s race”, (@toronto_hatecrimes_2024).
 
@@
 
-## Summary Statistics: Asian and South Asian Communities
+## Summary Statistics: East/Southeast Asian and South Asian Communities
 
-@fig-Asian-Bias-Summary presents summary statistics for hate crimes specifically targeting Asian racial groups.
-fig-East-Asn-Summary and fig-South-Asn-Summary show the increase in the 2 major Asian race groups being East/Southeast Asians and South Asians, respectively.
-The data reveals a steady increase in incidents against South Asians from 2018 to 2023, with notable spikes in crimes against East As Asians in particular.
-The increase in Asian biased crimes also show an increase in different offences and location of offence occurrences.
+@tbl-Asian-Bias-Summary presents summary statistics for hate crimes specifically targeting Asian racial groups.
+@tbl-East-Asn-Summary and @tbl-South-Asn-Summary show the increase in the 2 major Asian race groups being East/Southeast Asians and South Asians, respectively.
+The data reveals a steady increase in incidents against South Asians in 2021, with a notable spike in crimes against East Asians.
+The increase in Asian biased crimes also show an increase in different kinds of offences and location of hate crime occurrences.
 
 ```{r}
-#| label: fig-Asian-Bias-Summary
-#| fig-cap: Summary of Asian Biased Hate Crimes in Toronto
+#| label: tbl-Asian-Bias-Summary
+#| table-cap: Summary of Asian Biased Hate Crimes in Toronto
 #| echo: false
@@
 ```{r}
-#| label: fig-East-Asn-Summary
-#| fig-cap: Summary of increase in hate crimes against East/Southeast Asians
+#| label: tbl-East-Asn-Summary
+#| table-cap: Summary of increase in hate crimes against East/Southeast Asians
 #| echo: false
@@
 ```{r}
-#| label: fig-South-Asn-Summary
-#| fig-cap: Summary of increase in hate crimes against South Asians
+#| label: tbl-South-Asn-Summary
+#| table-cap: Summary of increase in hate crimes against South Asians
 #| echo: false
@@
 ```
+
+## Offence and location trends in Asian Biased Hate Crimes
+@fig-Asian-Offence-Types and @fig-Asian-Location-Types show spikes in assault and public based hate crimes in 2021 towards East/Southeast and South Asians, including racially mixed Asians.
 
@@
 
-Based on the above figures, there is evidence that
+Based on the above figures, there is evidence that the surge in hate crimes against East/Southeast and South Asians during 2021 correlates strongly with the social and political context surrounding the COVID-19 pandemic.
+Specifically a spike of over 1000% in hate crimes motivated by anti-East/Southeast Asian sentiment in @tbl-East-Asn-Summary.
+As the virus's origins were linked to Asia, widespread misinformation and xenophobia fueled anti-Asian sentiment.
+The sharp rise in hate crimes during this period is indicative of a broader global trend of increased racial targeting and discrimination, but concentrated within denser, diverse urban populations.
+While 2022 and 2023 saw a reduction in incidents, the numbers remain higher than pre-pandemic levels, suggesting that the effects of this wave of racism may have long-term impacts.
+Notably in @fig-Asian-Offence-Types and @fig-Asian-Location-Types, public settings such as businesses and streets became hotspots for these crimes, with assaults making up a significant portion of the incidents.
+In 2021, public assaults against Asians peaked and remained higher than pre-pandemic levels even into 2022 and 2023.
+While there was a reduction in incidents in these later years, the overall numbers still surpass those observed before the pandemic, suggesting long-lasting impacts of this wave of racism.
 
-The surge in hate crimes against East/Southeast Asians during 2020 and 2021 correlates strongly with the social and political context surrounding the COVID-19 pandemic.
-As the virus's origins were linked to Asia, widespread misinformation and xenophobia fueled anti-Asian sentiment.
-The sharp rise in hate crimes during this period is indicative of a broader global trend of increased racial targeting and discrimination.
-While 2022 and 2023 saw a reduction in incidents, the numbers remain higher than pre-pandemic levels, suggesting that the effects of this wave of racism may have long-term impacts.
-In contrast, the steady rise in hate crimes against South Asians points to a broader, underlying trend of racial bias that extends beyond the pandemic.
+In contrast, the steady rise since 2021 in @tbl-South-Asn-Summary in hate crimes against South Asians points to a broader, underlying trend of racial bias that extends beyond the pandemic.
 Anti-immigrant sentiment, exacerbated by policies affecting international students and immigration more generally, likely played a role in these trends.
-Despite government interventions aimed at supporting immigrant communities during the pandemic (Government of Canada, 2020), the persistent hostility faced by South Asians reflects the ongoing challenges of systemic racism in Canadian society.
-Meanwhile, the consistent levels of anti-Black hate crimes highlight ongoing systemic racism.
+Despite government interventions aimed at supporting immigrant communities during the pandemic (@gov_canada_2020_students), the persistent hostility faced by South Asians reflects the ongoing challenges of systemic racism in Canadian society.
+
+Meanwhile, the consistent levels of anti-Black hate crimes highlight ongoing systemic racism, highlighted in @fig-Race-Bias-Fig.
 Although hate crimes against Asians have decreased since their peak, the persistence of incidents reflects the long-term impact of these biases, necessitating continued policy focus on combating racial discrimination.
+
 Both the rise and persistence of these hate crimes underscore the importance of targeted policy responses.
@@
 # References
-
-City of Toronto. 2024. Hate Crimes Open Data. Published by Policy, Planning, Finance & Administration. Retrieved from https://www.toronto.ca/open-data-catalogue/hate-crimes-open-data.
-
-Gelfand, Sharla. 2022. Opendatatoronto: Access the City of Toronto Open Data Portal. Retrieved from https://CRAN.R-project.org/package=opendatatoronto.
-
-Government of Canada.
-2020.
-Measures to Support International Students During the COVID-19 Pandemic.
-Immigration, Refugees and Citizenship Canada.
-Retrieved from https://www.canada.ca/en/immigration-refugees-citizenship/services/coronavirus-covid19.html.
-
-Government of Canada.
-2023.
-Annual Report on the Operation of the Canadian Multiculturalism Act 2021-2022.
-Retrieved from https://www.canada.ca/en/canadian-heritage/corporate/publications/plans-reports/annual-report-canadian-multiculturalism-act-2021-2022.html.
-
-Hadley Wickham, Lionel Henry, Romain François. 2023. Dplyr: A Grammar of Data Manipulation. Retrieved from https://CRAN.R-project.org/package=dplyr.
-
-Kirill Müller, Romain Francois, Hadley Wickham. 2023. Tibble: Simple Data Frames. Retrieved from https://CRAN.R-project.org/package=tibble.
-
-The Link. 2022. Canada Has a Rampant South Asian Racism Problem. The Link Newspaper. Retrieved from https://thelinknewspaper.ca/article/canada-has-a-rampant-south-asian-racism-problem.
-
-R Core Team. 2023. R: A Language and Environment for Statistical Computing. Vienna, Austria: R Foundation for Statistical Computing. Retrieved from https://www.R-project.org/.
-
-Wickham, Hadley. 2016. Ggplot2: Elegant Graphics for Data Analysis. Springer-Verlag New York. Retrieved from https://ggplot2.tidyverse.org.
-
-Wickham, Hadley, Mara Averick, Jennifer Bryan, Winston Chang, Lucy D’Agostino McGowan, Romain François, Garrett Grolemund, et al.
-2019.
-“Welcome to the Tidyverse.” Journal of Open Source Software 4 (43): 1686.
-Retrieved from https://doi.org/10.21105/joss.01686.
-
-Xie, Yihui. 2023. Knitr: A General-Purpose Package for Dynamic Report Generation in R. Retrieved from https://yihui.org/knitr/.

### 2024-09-28 scripts/00-simulate_data.R
added small tests

@@ -23,5 +23,18 @@ simulated_data <- tibble(
   OFFENCE_GROUP = sample(cleaned_data$OFFENCE_GROUP, number_of_records, replace = TRUE)
 )
 
+## Tests ##
+
+# Check OCCURRENCE_YEAR range
+range(simulated_data$OCCURRENCE_YEAR)
+
+# Check duplicate rows
+sum(duplicated(simulated_data))
+
+# Check unique occurrences for RACE_BIAS, LOCATION_GROUP, and OFFENCE_GROUP for distribution
+table(simulated_data$RACE_BIAS)
+table(simulated_data$LOCATION_GROUP)
+table(simulated_data$OFFENCE_GROUP)
+
 #### Write_csv
 write_csv(data, file = "data/raw_data/simulated.csv")

### 2024-09-28 paper/paper.qmd
proper cross referencing

@@
 
-The following structure of this paper is as follows: Section @sec-mydatasection presents the data sources and methodology, followed by Section 3, which discusses the results and their implications.
-Section 4 concludes with a discussion of the limitations of the study and suggestions for future research.
+The following structure of this paper is as follows:  @sec-mydatasection presents the data sources and methodology, then  @sec-myresults followed by  @sec-mydisc, which discusses the results and their implications.
+@sec-mylimits concludes with a discussion of the limitations of the study and suggestions for future research.
 
@@
 
-The columns included in the analysis are Primary Offences (Regrouped), Race Bias and Occurance Year.
+The columns included in the analysis are Primary Offences (Regrouped), Race Bias and Occurence Year.
 Due to the analysis’ focus on Asian Bias hate crimes, races of only East, Southeast and South Asian descent were included.
@@
 
-# Results
+# Results {#sec-myresults}
 
@@
 
-# Discussion
+# Discussion {#sec-mydisc}
 
@@
 
-# Limitations and Next Steps
+# Limitations and Next Steps {#sec-mylimits}
 
